---
title: "ESM 244 Lab 9"
author: "Sara Orofino"
date: "3/7/2019"
output: html_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(beepr) # beep (on CRAN)
library(datapasta) # copy/paste (on CRAN)
library(tidyverse) 
library(igraph) # network stuff (on CRAN)
library(multiplex) # network stuff (on CRAN)
library(ggraph) # More graph stuff (on CRAN)
library(gt) # tables
library(ggalluvial) # Sankeys (on CRAN)
library(readxl) # Get .xls (on CRAN)
library(praise) # Praaaise! 
```

###Part 1. Network analysis examples - graphy theory (Les Mis character connections)

Get the data:

```{r}

lm_df <- read.gml("lesmis.gml")

#directed - not a direction in the graph theory; characters are connected but not in a directed way 

les_mis <- graph_from_data_frame(lm_df, directed = FALSE)  
beep(sound = 3) #intended to be an alarm for code chunks that take a long time to run

praise() # Give's awesome random encouragement to you and your group! 

```


Find some quantitative metrics:
```{r}

#Graph diameter:
diameter(les_mis) # Smallest maximum distance (links)

# Farthest vertices
farthest_vertices(les_mis)


```


Plot it: 
```{r}
# Alternatively, get the graph data as igraph right away with igraph::read_graph()
# les_mis <- read_graph("lesmis.gml", format = "gml")

plot(les_mis,
     vertex.color = "orange",
     vertex.frame.color = "NA",
     vertex.size = 5,
     vertex.label.color = "black",
     vertex.label.cex = 0.5,
     vertex.label.dist = 2)
```

A little more customized graphs using gg formats:  

```{r}

# geom_edge_links web format
# geom_node_text to show the nodes 

ggraph(les_mis, layout = "kk") +
  geom_edge_link() +
  geom_node_text(aes(label = name), size = 2, color = "white") +
  theme_dark()

# linear layout of characters with the half circle arcs connecting characters 

ggraph(les_mis, layout = "linear") +
  geom_edge_arc(alpha = 0.8) +
  geom_node_text(aes(label = name), angle = 90, size = 2, hjust = 1) +
  theme_graph()

# plotting chacacters as linear data coerced into levels

ggraph(les_mis, layout = "tree") +
  geom_edge_fan(color = "purple") +
  geom_node_text(aes(label = name), size = 2, color = "blue", angle = 45) +
  theme_void()

beep(3)
  
  

  
```

Or the ones that we're most used to seeing:  

```{r}

ggraph(les_mis, layout = "linear", circular = TRUE) +
  geom_edge_arc(alpha = 0.5) +
  geom_node_point(aes(color = name), show.legend = FALSE) +
  geom_node_text(aes(label = name), size = 3, hjust = "outward") +
  theme_void()

# Note you can show corrrelation using hierarchial clustering (correlation matrix - calculated when did dendrograms)

```

###Part 2. Sankey Diagrams  

- alluvial
- ggalluvial 
- NetworkD3
- ggforce (in ggplot - geom_parallel?)

```{r}


sdf <- read_csv("sankey_df.csv") %>% 
  select(-X1)


```


```{r}

ggplot(sdf, aes(y=weight, axis1 = before, axis2 = after)) +
  geom_alluvium(aes(fill = before, color = before), show.legend = FALSE) +
  geom_stratum(width = 1/5, color = "gray") +
  geom_text(stat = "stratum", label.strata = TRUE) +
  scale_fill_manual(values = c("purple","blue","green")) +
  scale_color_manual(values = c("purple","blue","green")) +
  scale_x_discrete(limits = c("Before", "After"), expand = c(0,0)) +
  theme_minimal()


```

###Part 3. A few of Allison's favorite things right now...

####3a. Creating your own tibbles:  

A tibble is like a data frame with a little more functionality. From r4ds (R for Data Science): "Tibbles are data frames, but they tweak some older behaviours to make life a little easier." Tibble and data frame are mostly used interchangeably.     

```{r}

# Create little data frames from scratch - good for testing out new code!
# In the first line use ~ to identify each of the column names
# The next lines will be all your row values (remember the number of values should match the number of columns)

my_tibble <- tribble(
  ~sara, ~made, ~this, ~awesomething,
  1, "yes", 0, 10000,
  2, "no", 4, 50000,
  3, "maybe", 10, 40000,
  4, "nah", 7, 35000
)

# And it works just like anything else:
ggplot(my_tibble, aes(x = sara, y = awesomething)) +
  geom_point(aes(color = made), size = 10) +
  scale_color_manual(values = c("red","orange","yellow", "blue")) +
  theme_dark()
```

Check out Jenny Bryan's awesome package 'reprex' - to make reproducible examples 


####3b. datapasta!

Can I copy and paste data from outside sources and have R store it as a data frame?   

```{r}
# tools --> addins --> browse addins --> paste as tribble 


weather_data <- tibble::tribble(
                                                  ~X,          ~Location, ~Min, ~Max,
                                    "Partly cloudy.",         "Brisbane",  19L,  29L,
                                    "Partly cloudy.", "Brisbane Airport",  18L,  27L,
                                  "Possible shower.",       "Beaudesert",  15L,  30L,
                                    "Partly cloudy.",        "Chermside",  17L,  29L,
                    "Shower or two. Possible storm.",           "Gatton",  15L,  32L,
                                  "Possible shower.",          "Ipswich",  15L,  30L,
                                    "Partly cloudy.",    "Logan Central",  18L,  29L,
                                     "Mostly sunny.",            "Manly",  20L,  26L,
                                    "Partly cloudy.",    "Mount Gravatt",  17L,  28L,
                                  "Possible shower.",            "Oxley",  17L,  30L,
                                    "Partly cloudy.",        "Redcliffe",  19L,  27L
                    )

weather_data <- rename(weather_data, Condition = X)

```


####3c. Beautiful customized tables with 'gt'

```{r}
weather_data %>% 
  gt() %>% 
  tab_header(
    title = "Check Out My Awesome Table",
    subtitle = "That's rad"
  ) %>% 
  fmt_number(
    columns = vars(Min),
    decimals = 2
  ) %>% 
  cols_move_to_start(
    columns = vars(Location)
  ) %>% 
  data_color(
    columns = vars(Min),
    colors = scales::col_numeric(
      palette = c("darkblue", "blue"),
      domain = c(14,20)
    )
  ) %>% 
  tab_options(
    table.background.color = "grey",
    heading.background.color = "black",
    column_labels.background.color = "green"
  )

# Check out the spanner column label function to add a bar over the variables (for shiny)

```

####3d. Reading in straight from a URL  

```{r}
# use read_delim and have to specify the type of deliminator (in this case vertical bars |, but could be comma)

nuclear <- read_delim("https://www.nrc.gov/reading-rm/doc-collections/event-status/reactor-status/PowerReactorStatusForLast365Days.txt", delim = "|", col_names = TRUE)


```

```{r}

 # Read in excel files 

pesticides <- read_xls("PesticideResidues.xls")

# Specify if you want to skip rows using skip = , useful if metadata is present in the dataset, specify if the first column after the skip will be the column headers using col_names = TRUE

pest2 <- read_xls("PesticideResidues.xls", skip = 2, col_names = TRUE)

pest <- pest2 %>% 
  janitor::clean_names()
```

Round 244 Labs with some data wrangling and visualization:  

```{r}
# Use separate() to get Oakland, Oakland,CA and Oakland, California to all show city as just Oakland

crops <- pest %>% 
  filter(commodity == "KALE", !is.na(grower_city)) %>% 
  separate(grower_city, c("grow_city", "grow_state"), sep = ",") %>% 
  separate(collection_site_city, 
           c("market_city", "market_state"), sep = ",") %>% 
  group_by(organic_commodity, grow_city, market_city) %>% 
  tally()

# Now make a Sankey Diagram

ggplot(crops, aes(y = n, axis1 = organic_commodity, axis2 = grow_city, axis3 = market_city)) +
  geom_alluvium(aes(fill = organic_commodity, color = organic_commodity),
                show.legend = FALSE) +
  geom_stratum() +
  geom_text(stat = "stratum", label.strata = TRUE, size = 2)
  
praise()
```


