---
title: "ESM 244 Lab 9 Key"
author: "Allison Horst"
date: "March 3, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Today: 

- Some network graphs
- Sankey diagrams
- Datapasta
- Nice tables with gt
- A code alarm clock
- Some other useful things

```{r, warning = FALSE, message = FALSE}

library(beepr) # beep
library(datapasta) # copy/paste
library(tidyverse) 
library(igraph) # network stuff
library(multiplex) # network stuff
library(ggraph) # More graph stuff
library(gt) # tables
library(ggalluvial) # Sankeys
library(readxl)


```


###Part 1. Network analysis example (Les Miserables character connections)

Get the data
```{r}

# Stored as a .gml file (graph modeling language) format supports network data
# Use multiplex::read.gml()

lm_df <- read.gml("lesmis.gml")

# Convert to graph (igraph) with nodes & edges
les_mis <- graph_from_data_frame(lm_df, directed = FALSE)
beep(sound = 2)

```

Find some quantitative metrics:
```{r}

#Graph diameter:
diameter(les_mis) # Smallest maximum distance (links)

# Farthest vertices
farthest_vertices(les_mis)


```


Plot it: 
```{r}
# Alternatively, get the graph data as igraph right away with igraph::read_graph()
# les_mis <- read_graph("lesmis.gml", format = "gml")

plot(les_mis,
     vertex.color = "orange",
     vertex.frame.color = "NA",
     vertex.size = 5,
     vertex.label.color = "black",
     vertex.label.cex = 0.5,
     vertex.label.dist = 2)
```

Other graphing options:
ggraph
ggnet2
plot
+ more...

A really cool resource for network plots in ggraph: <https://www.data-imaginist.com/2017/ggraph-introduction-layouts/>

Want to do it more in the style of ggplot? Use ggraph. 
```{r}

# Then plot it
ggraph(les_mis, layout = 'kk') +
  geom_edge_link() +
  geom_node_text(aes(label = name), size = 2, color = "white") +
  theme_dark()

# Or an arc graph: 
ggraph(les_mis, layout = "linear") + 
  geom_edge_arc(alpha = 0.8) + 
  geom_node_text(aes(label = name), angle = 90, size = 2, hjust = 1) +
  theme_graph()

# Try layout = "circle", "tree"
ggraph(les_mis, layout = "tree") +
  geom_edge_fan(color = "gray90") + 
  geom_node_text(aes(label = name), size = 2, color = "black", angle = 45) +
  theme_void()

```

One like the layouts you're probably most used to seeing: 
```{r, fig.height = 6, fig.width = 6}
# And a circular form of linear arcs:

ggraph(les_mis, layout = "linear", circular = TRUE) +
  geom_edge_arc(alpha = 0.5) +
  geom_node_point(aes(colour = name), show.legend = FALSE) +
  geom_node_text(aes(label = name), size = 3, hjust = "outward") +
  theme_void()
  

```

###Part 2. Sankey diagrams showing flows

Note: there are other options for creating Sankey diagrams. 
Other packages to look into for Sankeys: alluvial, ggalluvial, NetworkD3

Wikipedia: "Sankey diagrams are a specific type of flow diagram, in which the width of the arrows is shown proportionally to the flow quantity."

Let's just get some simple data that I've created: 
```{r}

sdf <- read_csv("sankey_df.csv") %>% 
  select(-X1)

# OK, now I want to make a diagram of those flows...

```

Then plot it with ggalluvial: 
A cool resource: <https://matthewdharris.com/2017/11/11/a-brief-diversion-into-static-alluvial-sankey-diagrams-in-r/>
```{r}

ggplot(sdf, aes(y = weight, axis1 = before, axis2 = after)) +
  geom_alluvium(aes(fill = before, color = before), show.legend = FALSE, width = 1/5) +
  geom_stratum(width = 1/5, color = "gray") +
  geom_text(stat = "stratum", label.strata = TRUE) +
  scale_fill_manual(values = c("purple","blue","green")) +
  scale_color_manual(values = c("purple","blue","green")) +
  scale_x_discrete(limits = c("Before", "After"), expand = c(0,0)) +
  theme_minimal()

```

###Part 3. A few of my favorite things right now

####3a. Let's make our own tibble! 

A tibble is like a data frame with a little more functionality. From r4ds: "Tibbles are data frames, but they tweak some older behaviours to make life a little easier." Tibble and data frame are mostly used interchangeably. 

```{r}

my_tibble <- tribble(
  ~allison, ~made, ~this, ~table,
  1,"yes",0,10000,
  2,"no",10, 20000,
  3,"maybe",20, 5000,
  4,"yes",15, 12000,
  5,"no",25, 18000
)

# Then it just works like anything else...
ggplot(my_tibble, aes(x = allison, y = table)) +
  geom_point(aes(color = made), size = 10) +
  scale_color_manual(values = c("red","orange","yellow")) +
  theme_dark()

```

####3b. But now, datapasta! 

Cool, so I can make my own data frames (tibbles) right here in R in a really intuitive way. But a lot of times there's data somewhere that we want in R, but it doesn't exist in a super useful format online. That's annoying. Luckily for us, there's *datapasta*!

*datapasta* from Miles McBain: <https://github.com/MilesMcBain/datapasta>

Once you install datapasta, then you should be able to find the datapasta addin (Tools > Addins > datapasta paste as tribble). 

Go to page on "How to datapasta:" <https://cran.r-project.org/web/packages/datapasta/vignettes/how-to-datapasta.html>

- copy table to clipboard
- go to Tools -> Addins -> datapasta:: paste as tribble
- ta da!

- Want it to be even easier to use? Create a keyboard shortcut for addins! (mine is shift + command + t)...they might want to see how to create those R shortcuts? 

```{r}

weather_data <- tibble::tribble(
                                ~X,          ~Location, ~Min, ~Max,
                  "Partly cloudy.",         "Brisbane",  19L,  29L,
                  "Partly cloudy.", "Brisbane Airport",  18L,  27L,
                "Possible shower.",       "Beaudesert",  15L,  30L,
                  "Partly cloudy.",        "Chermside",  17L,  29L,
  "Shower or two. Possible storm.",           "Gatton",  15L,  32L,
                "Possible shower.",          "Ipswich",  15L,  30L,
                  "Partly cloudy.",    "Logan Central",  18L,  29L,
                   "Mostly sunny.",            "Manly",  20L,  26L,
                  "Partly cloudy.",    "Mount Gravatt",  17L,  28L,
                "Possible shower.",            "Oxley",  17L,  30L,
                  "Partly cloudy.",        "Redcliffe",  19L,  27L
  )

weather_data <- rename(weather_data, Condition = X)

```

You can also do this from things like an Excel file. Open the PesticideResidues.xls file on your computer. Select a subset of the cells (don't include headers for this example). Then use datapasta to create a tibble from it in R (remember, my shortcut is Control + Shift + t): 

```{r}

pest_sub <- tibble::tribble(
          ~V1,                           ~V2,      ~V3,          ~V4,           ~V5,
  "06-Jan-14",                        "KALE", 2140001L, "California",   "Riverside",
  "06-Jan-14",                        "KALE", 2140001L, "California",   "Riverside",
  "06-Jan-14",                        "KALE", 2140001L, "California",   "Riverside",
  "06-Jan-14",  "STRAWBERRY (ALL OR UNSPEC)", 2140002L, "California", "Santa Maria",
  "06-Jan-14",  "STRAWBERRY (ALL OR UNSPEC)", 2140002L, "California", "Santa Maria",
  "06-Jan-14",  "STRAWBERRY (ALL OR UNSPEC)", 2140002L, "California", "Santa Maria",
  "06-Jan-14",  "STRAWBERRY (ALL OR UNSPEC)", 2140002L, "California", "Santa Maria",
  "06-Jan-14",  "STRAWBERRY (ALL OR UNSPEC)", 2140002L, "California", "Santa Maria",
  "06-Jan-14",      "ORANGE (ALL OR UNSPEC)", 2140003L, "California",       "Arvin",
  "06-Jan-14",      "ORANGE (ALL OR UNSPEC)", 2140003L, "California",       "Arvin",
  "06-Jan-14",      "ORANGE (ALL OR UNSPEC)", 2140003L, "California",       "Arvin",
  "06-Jan-14", "PEAR, ASIAN (ORIENTAL PEAR)", 2140004L, "California",   "Kingsburg",
  "06-Jan-14", "PEAR, ASIAN (ORIENTAL PEAR)", 2140004L, "California",   "Kingsburg",
  "06-Jan-14", "PEAR, ASIAN (ORIENTAL PEAR)", 2140004L, "California",   "Kingsburg",
  "06-Jan-14",                     "SPINACH", 2140005L, "California",    "Highland",
  "06-Jan-14",                     "SPINACH", 2140005L, "California",    "Highland",
  "06-Jan-14",                     "SPINACH", 2140005L, "California",    "Highland"
  )

```

####3c. Beautiful tables with gt

The *gt* package by Rich Iannone is awesome (<https://github.com/rstudio/gt>). 

Cool, so *gt* just works...
```{r}
weather_data %>% 
  gt()
```

Or make it way better: 

<https://gt.rstudio.com/reference/tab_options.html>

```{r}
weather_data %>% 
  gt() %>% 
  tab_header(
    title = "An amazing title", # Add a title
    subtitle = "(by Allison!)"# And a subtitle
  ) %>%
  fmt_passthrough( # Not sure about this but it works...
    columns = vars(Location) # First column: supp (character)
  ) %>% 
  fmt_number(
    columns = vars(Min), # Second column: mean_len (numeric)
    decimals = 1 # With 4 decimal places
  ) %>% 
    fmt_number(
    columns = vars(Max), # Third column: dose (numeric)
    decimals = 1 # With 2 decimal places
  ) %>% 
  cols_move_to_start(
    columns = vars(Location)
  ) %>% 
  data_color( # Update cell colors...
    columns = vars(Min), # ...for mean_len column
    colors = scales::col_numeric(
      palette = c(
        "darkblue", "blue"), # Overboard colors! 
      domain = c(20,14) # Column scale endpoints
  )
  ) %>% 
  data_color(
    columns = vars(Max),
    colors = scales::col_numeric(
      palette = c(
       "yellow", "orange", "red"),
      domain = c(26,32)
      )
    ) %>% 
  tab_options(
    table.background.color = "purple",
    heading.background.color = "black",
    column_labels.background.color = "gray",
    heading.border.bottom.color = "yellow",
    column_labels.font.size = 10
  )

# And see https://gt.rstudio.com/reference/tab_options.html for mor overall formatting!
```

See more examples here: 
<https://github.com/allisonhorst/gt-awesome-tables>


###3d. Reading in straight from a URL

Google then direct to get to the UCI Machine Learning data for auto-mpg. Then read it in via the URL. 
Make sure to ask yourself: why is this convenient but risky? 

- Google: U.S. Nuclear Power Reactor Plant Status 
- From: data.gov
- Then click on 'text file' under Downloads and Resources
- Copy URL

Note: don't want to have reading from URL in document that's knitting (requires connection)...? Depends.
```{r}
# nuclear <- read_delim("https://www.nrc.gov/reading-rm/doc-collections/event-status/reactor-status/PowerReactorStatusForLast365Days.txt", delim = "|", col_names = TRUE)

# Because you always want a version of what you were working with, I recommend writing that to a csv (or a text file, or whatever)

# write_csv(nuclear, "nuclear.csv")

```

But that also works if it's in your working directory...

###3e. Reading in Excel files (and skipping top rows metadata)

Easy to read in Excel files (respecting the classes):
```{r}

pesticides <- read_xls("PesticideResidues.xls")

# Dang, there are two extra rows with information in them. Bummer. But: 

pest2 <- read_xls("PesticideResidues.xls", skip = 2, col_names = TRUE) # Ta da! 

pest <- pest2 %>% 
  janitor::clean_names()

```

Let's clean that up, do some wrangling, and a cool visualization to close out 244 lab for the year

Some wrangling of the pesticide data: 

```{r}

crops <- pest %>% 
  filter(commodity == "KALE",!is.na(grower_city)) %>% 
  separate(grower_city, c("grow_city","grow_state"), sep = ",") %>% 
  separate(collection_site_city, c("market_city","market_state"), sep = ",") %>% 
  group_by(organic_commodity, grow_city, market_city) %>% 
  tally()

```

Then a Sankey diagram! 
```{r}

ggplot(crops, aes(y = n, axis1 = organic_commodity, axis2 = grow_city, axis3 = market_city)) +
  geom_alluvium(aes(fill = organic_commodity, color = organic_commodity), show.legend = FALSE, width = 1/5) +
  geom_stratum(width = 1/5, color = "gray") +
  geom_text(stat = "stratum", label.strata = TRUE, size = 2) +
  scale_fill_manual(values = c("purple","blue")) +
  scale_color_manual(values = c("purple","blue")) +
  scale_x_discrete(limits = c("status", "grown","market"), expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  theme_bw()

```

Other things to show during this lab: 

- Comment out multiple lines with Command + Shift + c
- Selecting a whole line or parts: Command + Shift + arrow
- Just shift arrow for by letter

